---
title: "Real-time theming with bslib and thematic"
output: 
  html_document:
    code_folding: show
    theme:
      color-contrast-warnings: false
      bg: "#202123"
      fg: "#B8BCC2"
      primary: "#EA80FC"
      secondary: "#00DAC6"
      base_font:
        google: Prompt
      heading_font:
        google: Proza Libre
---

#```{r setup, include=FALSE}
#install.packages("remotes")
#install.packages("nflverse")
#remotes::install_github("FantasyFootballAnalytics/ffanalytics")
#install.packages(c("reshape", "MASS", "psych", "XML"), dependencies=TRUE)
#install.packages('ffsimulator', repos = c('https://ffverse.r-universe.dev', 'https://cloud.r-project.org'))
#install.packages('ffopportunity', repos = c('https://ffverse.r-universe.dev', 'https://cloud.r-project.org'))
#```

```{r setup, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(nflverse)
library(ffanalytics)
library(data.table)
library(purrr)
library(knitr)
library(writexl)
library(reshape)
library(MASS)
library(psych)
library(XML)
library(nflreadr)
library(stringr)
library(nflplotR)
library(gt)
```


## Scraping Projections
```{r}
raw_scrape <- scrape_data(
  src = c("CBS", "ESPN", "FantasyPros", "FantasySharks", "FFToday",
          "FleaFlicker", "FantasyFootballNerd", "NFL",
          "RTSports", "Walterfootball", "NumberFire"),
          pos = c("QB", "RB", "WR", "TE", "DST"),
          season = 2026, week = 0
)
```

## Seperating positions from scrape

```{r}
# This is the example from FFAnalytics
# It creates a projections table without the raw stats, which I can't use my custom league scoring for
# But, I still need it for things like ADP, ECR, AAV, and Uncertainty
# Which will be merged with the clean custom projections later
raw_projections <- projections_table(raw_scrape)

raw_projections <- raw_projections |> 
  add_ecr() |> 
  add_adp() |> 
  add_aav() |> 
  add_uncertainty()

raw_projections <- raw_projections |> 
  add_player_info()
```

```{r}
# Creating a function to clean the player names to avoid mismatches later
clean_player_name <- function(name_vec) {
  name_vec |> 
    # 1. Flip "Last, First" to "First Last"
    str_replace("^([^,]+),\\s*(.*)$", "\\2 \\1") |> 
    # 2. Remove common suffixes that break exact matching
    str_remove_all("\\b(Jr\\.?|Sr\\.?|III|II|IV|V)\\b") |> 
    # 3. Remove periods (converts "A.J." or "D.J." to "AJ" or "DJ")
    str_remove_all("\\.") |> 
    # 4. Clean up any accidental double spaces created by the removals
    str_squish()
}
# Taking the list of scraped data, and creating a data.frame
raw_scrape_data <- bind_rows(raw_scrape) |> 
  mutate(
    player_clean = clean_player_name(player)
  )
# Now we move the player_clean variable to a position that makes sense

raw_scrape_data <- raw_scrape_data |> 
  relocate(player_clean, .after = player)
```

```{r}
# that new data frame now being split up by position except kickers.  Fuck'em that's why.
qb_raw <- dplyr::filter(raw_scrape_data, pos == "QB") |> 
  dplyr::select(player_clean:fumbles_lost, data_src, id)

rb_raw <- dplyr::filter(raw_scrape_data, pos == "RB") |>
  dplyr::select(player_clean:games, rush_att:fumbles_lost, data_src, id, rec_tgt:rec_tds)

wr_raw <- dplyr::filter(raw_scrape_data, pos == "WR") |> 
  dplyr::select(player_clean:games, rush_att:fumbles_lost, data_src, id, rec_tgt:rec_tds)

te_raw <- dplyr::filter(raw_scrape_data, pos == "TE") |> 
  dplyr::select(player_clean:games, rush_att:fumbles_lost, data_src, id, rec_tgt:rec_tds)
  
dst_raw <- dplyr::filter(raw_scrape_data, pos == "DST" | pos == "D") |> 
  dplyr::select(player_clean:team, dst_safety:dst_ret_tds)
```

## Scoring the raw data we just finished making
```{r}
# Creating the custom scoring for each /pos/_raw
score_qb <- function(df){
  df |> 
    mutate(
      across(where(is.double), \(x) replace_na(x, 0)), # Changing N/A's to 0's because letters don't math well

      pass_pts = (pass_yds * 0.04) + (pass_tds *4) - (pass_int * 2),
      rush_pts = (rush_att * 0.3) + (rush_yds * 0.1) + (rush_tds * 6),
      fumble_pts = (fumbles_lost * -2),

      custom_fp = pass_pts + rush_pts + fumble_pts
    )
}

score_skill_pos <- function(df){
  df |> 
    mutate(
      across(where(is.double), \(x) replace_na(x, 0)),

      rush_pts = (rush_att * 0.3) + (rush_yds * 0.1) + (rush_tds * 6),
      rec_pts = (rec * 1) + (rec_yds * 0.1) + (rec_tds * 6),
      fumble_pts = (fumbles_lost * -2),

      custom_fp = rush_pts + rec_pts + fumble_pts
    )
}

# Running raw stats through custom scoring
qb_scored <- qb_raw |> score_qb()
rb_scored <- rb_raw |> score_skill_pos()
wr_scored <- wr_raw |> score_skill_pos()
te_scored <- te_raw |> score_skill_pos()
```

## Taking the scored data, and making it smush together
```{r}
qb_final_projections <- qb_scored |> 
  dplyr::filter_out(is.na(id)) |> 
  dplyr::group_by(id) |> 
  dplyr::summarise(
    player = first(player_clean),
    proj_fp_avg = mean(custom_fp, na.rm = TRUE),
    proj_fp_median = median(custom_fp, na.rm = TRUE),

    ## calculate std. devation and the number of sources per player
    proj_sd = sd(custom_fp, na.rm = TRUE), 
    num_sources = n_distinct(data_src), 
    .groups = "drop"
  )


rb_final_projections <- rb_scored |> 
  dplyr::filter_out(is.na(id)) |> 
  dplyr::group_by(id) |> 
  dplyr::summarise(
    player = first(player_clean),
    proj_fp_avg = mean(custom_fp, na.rm = TRUE),
    proj_fp_median = median(custom_fp, na.rm = TRUE),

    ## calculate std. devation and the number of sources per player
    proj_sd = sd(custom_fp, na.rm = TRUE), 
    num_sources = n_distinct(data_src), 
    .groups = "drop"
  )

wr_final_projections <- wr_scored |> 
  dplyr::filter_out(is.na(id)) |> 
  dplyr::group_by(id) |> 
  dplyr::summarise(
    player = first(player_clean),
    proj_fp_avg = mean(custom_fp, na.rm = TRUE),
    proj_fp_median = median(custom_fp, na.rm = TRUE),

    ## calculate std. devation and the number of sources per player
    proj_sd = sd(custom_fp, na.rm = TRUE), 
    num_sources = n_distinct(data_src), 
    .groups = "drop"
  )

te_final_projections <- te_scored |> 
  dplyr::filter_out(is.na(id)) |> 
  dplyr::group_by(id) |> 
  dplyr::summarise(
    player = first(player_clean),
    proj_fp_avg = mean(custom_fp, na.rm = TRUE),
    proj_fp_median = median(custom_fp, na.rm = TRUE),

    ## calculate std. devation and the number of sources per player
    proj_sd = sd(custom_fp, na.rm = TRUE), 
    num_sources = n_distinct(data_src), 
    .groups = "drop"
  )


# Now we arranged them neatly by averages, not because that's better,
# but because FFAnalytics doesn't produce a median number in their projections table command
qb_final_projections <- qb_final_projections |> 
  dplyr::arrange(desc(proj_fp_avg))

rb_final_projections <- rb_final_projections |> 
  dplyr::arrange(desc(proj_fp_avg))

wr_final_projections <- wr_final_projections |> 
  dplyr::arrange(desc(proj_fp_avg))

te_final_projections <- te_final_projections |> 
  dplyr::arrange(desc(proj_fp_avg))
```

## Righty ho, now, that's the boring and tedious parts done
## Now, we find our base lines for VOR, Calculate, and Bob's your uncle, and your dad

```{r}
# Let's get functional
calculate_VOR <- function(df, replacement_rank) {
  replacement_points <- df |> 
    arrange(desc(proj_fp_avg)) |> 
    slice(replacement_rank) |> 
    pull(proj_fp_avg)

  df |> 
    arrange(desc(proj_fp_avg)) |> 
    mutate(
      projection_rank = row_number(),
      replacement_points = replacement_points,
      vor = proj_fp_avg - replacement_points,
      vor_rank = min_rank(desc(vor))
    )
}

# Staying with the seperate pieces theme, now we find each position VOR
qb_vor <- qb_final_projections |> 
  calculate_VOR(replacement_rank = 13)

rb_vor <- rb_final_projections |> 
  calculate_VOR(replacement_rank = 37)

wr_vor <- wr_final_projections |> 
  calculate_VOR(replacement_rank = 41)

te_vor <- te_final_projections |> 
  calculate_VOR(replacement_rank = 13)
```

## This is where the wheel completes it's lap.  We are going to use that FFAnalytics example and attach data from it here

```{r}
# Attaching ADP, ECR, AAV, and Uncertainty
# Getting Functional first
attach_market_data <- function(final_df, raw_proj) {
  final_df |> 
    left_join(
      raw_proj |> 
        dplyr::select(id, overall_ecr, adp, adp_sd, aav, uncertainty, tier, team) |> 
        distinct(id, .keep_all = TRUE), 
      by = "id"
    ) |> 
      relocate(team, overall_ecr, adp, adp_sd, uncertainty, tier, .after = player)
}
qb_vor <- attach_market_data(qb_vor, raw_projections)
rb_vor <- attach_market_data(rb_vor, raw_projections)
wr_vor <- attach_market_data(wr_vor, raw_projections)
te_vor <- attach_market_data(te_vor, raw_projections)
```

## Now for the juice that really makes this worth the squeeze. Real Historical Stats.
```{r}
# pull 10 years of historical data.  Thank you NFLverse
historical_raw <- nflreadr::load_player_stats(seasons = 2015:2025, summary_level = "reg")

# Use the same scoring as earlier.  Consistency, thy name is boredom.
historical_scored <- historical_raw |> 
  filter(position %in% c("QB", "RB", "WR", "TE")) |> 
  mutate(
    across(where(is.numeric), \(x) replace_na(x, 0)),

    pass_pts = (passing_yards * 0.04) + (passing_tds * 4) - (passing_interceptions * 2),
    rush_pts = (carries * 0.3) + (rushing_yards * 0.1) + (rushing_tds * 6),
    rec_pts  = (receptions * 1.0) + (receiving_yards * 0.1) + (receiving_tds * 6),
    fumb_pts = (sack_fumbles_lost + rushing_fumbles_lost + receiving_fumbles_lost) * -2,
    
    custom_fp = pass_pts + rush_pts + rec_pts + fumb_pts
  )

# NFLverse uses different ID's & player_name ideal than FFA.  We clean that first, then find the total number of seasons per player...
historical_summary <- historical_scored |> 
  mutate(
    player_clean = clean_player_name(player_display_name)
  ) |> 
  group_by(player_clean, position) |> 
  summarise(
    seasons_played = n_distinct(season),
    career_games =  sum(games, na.rm = TRUE),
    hist_fp_avg = mean(custom_fp, na.rm = TRUE),
    hist_fp_max = max(custom_fp, na.rm = TRUE), 
    last_season_fp = custom_fp[which.max(season)],
    .groups = "drop"
  )
# ...And we average them out across the board.  Boredom thy name is consistency

# Okay, almost done, the next step is joinging the historical stuff with the positional VOR stuff
attach_historical_delta <- function(vor_df, hist_summary){
  vor_df |> 
    left_join(
      hist_summary |> dplyr::select(!(position)),
      by = c("player" = "player_clean")
    ) |> 
      mutate(
        hist_delta_avg = proj_fp_avg - hist_fp_avg,
        hist_delta_last = proj_fp_avg - last_season_fp,
        player_type = case_when(
          is.na(seasons_played) ~ "Rookie/No History",
          seasons_played <= 2 ~ "Young (1-2yrs)",
          TRUE ~ "Veteran (3+ yrs)"
        )
      ) |> 
        relocate(player_type, seasons_played, hist_fp_avg, hist_delta_avg, hist_delta_last, .after = proj_fp_avg)
}

# And then we do some more smushing
qb_vor <- attach_historical_delta(qb_vor, historical_summary)
rb_vor <- attach_historical_delta(rb_vor, historical_summary)
wr_vor <- attach_historical_delta(wr_vor, historical_summary)
te_vor <- attach_historical_delta(te_vor, historical_summary)
```


## Output to Excel for that nitty gritty stuff

```{r}
# Combine top starters by position
top_starters <- bind_rows(
  qb_vor |> dplyr::slice_head(n = 12) |> dplyr::mutate(pos = "QB"),
  rb_vor |> dplyr::slice_head(n = 24) |> dplyr::mutate(pos = "RB"),
  wr_vor |> dplyr::slice_head(n = 30) |> dplyr::mutate(pos = "WR"),
  te_vor |> dplyr::slice_head(n = 12) |> dplyr::mutate(pos = "TE")
)

 write.csv(top_starters, file = "custom_projections.csv")

# Export Multi-Tab Excel Spreadsheet (v1)
# writexl::write_xlsx(
#  list(
#    "Top Starters" = top_starters,
#    "QBs"          = qb_vor,
#    "RBs"          = rb_vor,
#    "WRs"          = wr_vor,
#    "TEs"          = te_vor
#  ),
#  path = "fantasy_projections_v1.xlsx"
#)
```






## Plotting some things

```{r}
# Cheat Sheet
# Build the formatted table
gt_cheat_sheet <- top_starters |> 
  dplyr::select(
    pos, projection_rank, player, team, proj_fp_avg, vor, adp, overall_ecr, uncertainty
  ) |> 
  gt(groupname_col = "pos") |> 
  tab_header(
    title = md("**2026 Fantasy Football Draft Cheat Sheet**"),
    subtitle = "Custom Point-Per-Carry & PPR Scoring | Value Over Replacement (VOR) Analysis"
  ) |> 
  # Automatically converts team abbreviations into official NFL logos
  gt_nfl_logos(columns = team) |> 
  fmt_number(columns = c(proj_fp_avg, vor, adp, overall_ecr), decimals = 1) |> 
  data_color(
    columns = vor,
    palette = c("#F1F8E9", "#A5D6A7", "#2E7D32") # Light to dark green highlighting
  ) |> 
  cols_label(
    projection_rank = "Pos Rank",
    player          = "Player",
    team            = "Team",
    proj_fp_avg     = "Proj FP",
    vor             = "VOR",
    adp             = "ADP",
    overall_ecr     = "ECR",
    uncertainty     = "Risk Index"
  ) |> 
  tab_options(
    table.font.size           = px(13),
    heading.title.font.size   = px(18),
    column_labels.font.weight = "bold",
    row_group.font.weight     = "bold"
  )

# Preview in RStudio viewer or save as HTML/PNG to email/text
gt::gtsave(gt_cheat_sheet, filename = "fantasy_cheat_sheet_v1.html")
# Optional: save as PNG image (requires webshot2 package)
# gt::gtsave(gt_cheat_sheet, filename = "fantasy_cheat_sheet_v1.png")

# Filter to top tier starters for a clean visual (e.g. top 8 per position)
plot_data <- top_starters |> 
  group_by(pos) |> 
  slice_head(n = 8) |> 
  ungroup() |> 
  mutate(
    # Order players so highest VOR renders at the top
    player_order = reorder(paste0(player, " (", pos, ")"), vor)
  )

# Column Plot
vor_plot <- ggplot(plot_data, aes(x = vor, y = player_order, fill = pos)) +
  geom_col(width = 0.7, show.legend = FALSE) +
  facet_wrap(~ pos, scales = "free_y", ncol = 2) +
  theme_minimal(base_size = 12) +
  labs(
    title = "Top Starter Value Over Replacement (VOR)",
    subtitle = "Comparing Custom Projected Value Across Positions",
    x = "Value Over Replacement (Points)",
    y = NULL
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    strip.text = element_text(face = "bold", size = 13),
    panel.grid.minor = element_blank()
  )

# Display plot
print(vor_plot)

# Save image file to send via message
ggsave("vor_top_starters_v1.png", vor_plot, width = 10, height = 8, dpi = 300)

# Cheat Sheet v2
library(gt)
library(nflplotR)
library(htmltools)
library(dplyr)

# -------------------------------------------------------------------------
# 1. Helper Function: Builds an Ultra-Compact GT Table for One Position
# -------------------------------------------------------------------------
create_compact_pos_table <- function(df, pos_name, header_color, limit_n = 25) {
  
  df |> 
    dplyr::slice_head(n = limit_n) |> 
    dplyr::select(projection_rank, player, team, proj_fp_avg, vor, adp) |> 
    gt() |> 
    tab_header(
      title = md(paste0("**", pos_name, "**"))
    ) |> 
    # Small NFL logos
    gt_nfl_logos(columns = team, height = 18) |> 
    fmt_number(columns = c(proj_fp_avg, vor, adp), decimals = 1) |> 
    # Subtle VOR Heatmap
    data_color(
      columns = vor,
      palette = c("#F1F8E9", "#C8E6C9", "#81C784", "#388E3C"),
      alpha = 0.85
    ) |> 
    cols_label(
      projection_rank = "#",
      player          = "Player",
      team            = "",
      proj_fp_avg     = "FP",
      vor             = "VOR",
      adp             = "ADP"
    ) |> 
    tab_options(
      # Compact padding & font settings to remove vertical bloat
      table.font.size                   = px(11),
      heading.title.font.size           = px(14),
      heading.background.color          = header_color,
      column_labels.font.size           = px(10),
      column_labels.font.weight         = "bold",
      column_labels.padding             = px(3),
      data_row.padding                  = px(2),  # Tight row spacing
      table.width                       = pct(100),
      table.border.top.style            = "none",
      table.border.bottom.style         = "none"
    )
}

# -------------------------------------------------------------------------
# 2. Build Individual Cards
# -------------------------------------------------------------------------
# Set custom row limits (e.g., Top 12 QB, Top 24 RB, Top 30 WR, Top 12 TE)
tbl_qb <- create_compact_pos_table(qb_vor, "Quarterbacks", "#2B4C7E", limit_n = 12)
tbl_rb <- create_compact_pos_table(rb_vor, "Running Backs", "#1B5E20", limit_n = 24)
tbl_wr <- create_compact_pos_table(wr_vor, "Wide Receivers", "#B71C1C", limit_n = 30)
tbl_te <- create_compact_pos_table(te_vor, "Tight Ends",     "#E65100", limit_n = 12)

# -------------------------------------------------------------------------
# 3. Assemble into a Side-by-Side Flexbox Dashboard
# -------------------------------------------------------------------------
draft_dashboard <- htmltools::tags$div(
  style = "font-family: -apple-system, BlinkMacSystemFont, 'Segoe UI', Roboto, Helvetica, Arial, sans-serif; background-color: #F8F9FA; padding: 15px;",
  
  # Dashboard Header
  htmltools::tags$div(
    style = "margin-bottom: 15px; text-align: center;",
    htmltools::tags$h2(style = "margin: 0; color: #1A1A1A; font-size: 22px;", "2026 DRAFT DAY VALUE BOARD"),
    htmltools::tags$p(style = "margin: 4px 0 0 0; color: #666; font-size: 12px;", "Custom PPR + 0.3 PPA | Value Over Replacement (VOR) Tracker")
  ),
  
  # Responsive Flex Grid (2x2 or 4-Column Layout)
  htmltools::tags$div(
    style = "display: flex; flex-wrap: wrap; gap: 12px; justify-content: center; align-items: flex-start;",
    
    # QB Card
    htmltools::tags$div(style = "flex: 1 1 280px; max-width: 340px; background: white; box-shadow: 0 2px 5px rgba(0,0,0,0.08); border-radius: 6px; overflow: hidden;", as_raw_html(tbl_qb)),
    
    # RB Card
    htmltools::tags$div(style = "flex: 1 1 280px; max-width: 340px; background: white; box-shadow: 0 2px 5px rgba(0,0,0,0.08); border-radius: 6px; overflow: hidden;", as_raw_html(tbl_rb)),
    
    # WR Card
    htmltools::tags$div(style = "flex: 1 1 280px; max-width: 340px; background: white; box-shadow: 0 2px 5px rgba(0,0,0,0.08); border-radius: 6px; overflow: hidden;", as_raw_html(tbl_wr)),
    
    # TE Card
    htmltools::tags$div(style = "flex: 1 1 280px; max-width: 340px; background: white; box-shadow: 0 2px 5px rgba(0,0,0,0.08); border-radius: 6px; overflow: hidden;", as_raw_html(tbl_te))
  )
)

# -------------------------------------------------------------------------
# 4. Save HTML File
# -------------------------------------------------------------------------
htmltools::save_html(draft_dashboard, file = "draft_day_cheat_sheet.html")
```